# Starting Data Analysis

## Creating A Project

You should always use the project functonality when you work with R and RStudio.

This also sets the working-directory.

```{r}
getwd()
```


## Folder Structure


> Show new and existing directory options. Some people may already have collected data.

    +---Data
    |
    +---Output
    

## Packages

> We can prime this by telling people about the R community in the introduction
    
installing and loading (maybe show updating in RStudio)    
    
## Importing Data    

> This is rather short and not worth a chapter on it's own.

Filetypes:

* csv
* xlsx


```{r}
library(readr)
```


Here we introduce the `readr` and `readxl` packagtes.


```{r include=FALSE}
write_csv(iris, path = "iris.csv")
```



```{r readr-1}
my_iris <- read_csv("iris.csv")


head(my_iris, n = 5)
```


this is useful because we can fix the column types like this.


```{r}
my_iris <- read_csv(
    file = "iris.csv",
    col_types = cols(
      Sepal.Length = col_double(),
      Sepal.Width = col_double(),
      Petal.Length = col_double(),
      Petal.Width = col_double(),
      Species = col_character()
    )
)

```

