# Next Steps

## How To Get Help

* Stack overflow

## More on packages

* what goes here?

## Writting clean code

1. Do not exceed 80 characters per line
2. Indent 2 spaces (this should be the default setting for tab to spaces)

You should always use underscores as separators in variable names. 

### Variable Names

```{r}
avg_length <- 3
avg_len <- 4

#Bad 
avgLength <- 5

```


you can use numbers in variable names but you cannot start a variable name with a number. 
Most of the time you feels like using a number in a variable name you should consider using
a list instead.

```{r}
value_1 <- c(27, 19)
value_2 <- c(13, 42)

# Better
values <- list(c(27, 19), c(13, 42))
```


### Assignment Operator

Always use `<-` for assignment and `=` for named arguments in function calls.

```{r}
# good
x <- 10

# bad
x = 10

# good
z <- t.test(1:10, y = 10:1, paired = T)
```

### Spaces

* Never use a space before a comma, always use a space after a comma.
* always put a space before and after a binary operator `+`, `-`, `*`, `=`, `==`, etc.

### Long Lists 

```{r}
# good
students <- list(
  jon = "Jon Snow",
  greg = "Greg Martin",
  reto = "Reto Wyss",
  severus = "Severus Snape",
  barney = "Barney Stinson"
)

students <- list(
  jon     = "Jon Snow",
  greg    = "Greg Martin",
  reto    = "Reto Wyss",
  severus = "Severus Snape",
  barney  = "Barney Stinson"
)

# not our preference but kind of ok
students <- list(jon = "Jon Snow",
                 greg = "Greg Martin", 
                 reto = "Reto Wyss",
                 severus = "Severus Snape",
                 barney = "Barney Stinson")

# bad
students <- list(jon = "Jon Snow", greg = "Greg Martin", reto = "Reto Wyss", severus = "Severus Snape", barney = "Barney Stinson")
```


### Function Calls


```{r}
# good
x <- mean(1:10)

# (1) piped unnamed argument
better_iris <- iris %>% 
  mutate(
    petal_volume = Petal.Length * Petal.Width / 2 * pi,
    sepal_volume = Sepal.Length * Sepal.Width / 2 * pi
  )

# (2) unnamed argument by itself, named arguments grouped
better_iris <- mutate(iris,
  petal_volume = Petal.Length * Petal.Width / 2 * pi,
  sepal_volume = Sepal.Length * Sepal.Width / 2 * pi
)

# (3) ok
better_iris <- iris %>% 
  mutate(petal_volume = Petal.Length * Petal.Width / 2 * pi,
         sepal_volume = Sepal.Length * Sepal.Width / 2 * pi)
```


### Function Definition

```{r}
# good
my_function <- function(argument_one, argument_two) {
  # code here
}

# bad
my_function <- function(argument_one, argument_two)
{
  # code here
}


# good
my_function <- function(argument_one, 
                        argument_two, 
                        argument_three, 
                        argument_four) {
  # code here
}
```


### Exercises

Here you find some code snippets, judge for yourself whether these are written in good style.

* We can reference some stuff here

## What else you can do with R

* dashboards
* blogdown
* deeplearning
 


